Binning Strategies and Related Loss for Large Data

### 2015-02-17 Paper.Rnw
fixed?

qplot(data=d1.hex, x=G, y=SO, geom="point", size=Freq) + scale_size_area("Frequency", trans="log", breaks=c(1, 10, 100, 1000, 10000))+ theme_bw() + xlab("Games (Count)")+ ylab("Strike Outs (Count)") + theme(legend.position="bottom")

@@

      temp <- data.frame(   
                x=c(-20, -10,   0,  10,  20,  30,  40, -20, -20, -20, -20, -20, -20), 
             xend=c(-20, -10,   0,  10,  20,  30,  40,  40,  40,  40,  40,  40,  40), 
                y=c(  0,   0,   0,   0,   0,   0,   0,   0,  10,  20,  30,  40,  50), 
             yend=c( 50,  50,  50,  50,  50,  50,  50,   0,  10,  20,  30,  40,  50), 
            group=c(  1,   2,   3,   4,   5,   6,   7,   8,   9,  10,  11,  12,  13))
      qplot(data=d[,], x=x, y=y, alpha=a,  geom="point", 
            xlim=c(-20, 40), ylim=c(0, 50)) +
        annotate("polygon", x=c(10, 10, 20, 20), y=c(20, 30, 30, 20), 
                 fill="blue", alpha=.25) +
        annotate("polygon", x=c(0, 0, 10, 10), y=c(20, 30, 30, 20), 
                 fill="green", alpha=.25) +
        scale_alpha(range=c(.5, 1), guide=FALSE) +
        scale_colour_identity(guide=FALSE) + 
        scale_shape(solid=T,guide=FALSE) +
        theme_bw() +
        annotate("point", x=rep(seq(-15, 35, 10),each=7), 
               y=rep(seq(-5, 55, 10), times=6), 
               shape='+', size=8, colour="gray90") + 
        geom_point(data=d[which(!is.na(d$xbin)),], 
                   aes(x=x, y=y, shape=factor(xbin), colour=factor(xbin)), 
                   size=I(3), aes.inherit=FALSE) + 
        scale_shape_manual(name="Bin", values=c("5"=16, "15"=15), 
                           labels=c("(5,25)", "(15, 25)"),guide=FALSE) +
        scale_colour_manual(values=c("5"="green", "15"="blue"), name="Bin", 
                            labels=c("(5,25)", "(15, 25)"),guide=FALSE)+
        annotate("segment", x=temp$x, y=temp$y, 
                 xend=temp$xend, yend=temp$yend, group=temp$group)  + ylim(c(0, 50))
      ggsave("./figure/Binning-DataAndBin.pdf", width=5, height=5, units="in")

@@

\begin{figure}[hbtp]
  \centering
  \subfloat[Standard Algorithm]{\includegraphics[keepaspectratio=true, width=3in]{./figure/Binning-DataAndBin.pdf}}
%   \subfloat{\includegraphics[keepaspectratio=TRUE, width=2in]
%   {./figure/RandomBinningDataPoints1}}
  \subfloat[Random Algorithm]{\includegraphics[keepaspectratio=TRUE, width=3in]
  {./figure/RandomBinningDataPoints1.pdf}}
  \caption{Points are assigned to bins centered at (5, 25) an (15,25). Points in the standard algorithm are contained within the defined bin region, where points from the random algorith may be assigned to bins from outside the defined bin regions.}
  \label{fig:BinningTutorial}
\end{figure}

@@

\section{Discussion and Examples}
\subsection{Comparison of Loss in Binning Algorithms}
<<LossTableGraphs,echo=F,results="hide",cache=TRUE, eval=TRUE>>=
# bins <- rbind(c(1,1), c(2,2), c(2,5), c(5, 2), c(4,4), c(3, 5), c(5, 3), c(5, 5), c(5, 25), c(25, 5), c(2, 50), c(50, 2), c(5, 50))
computeLoss <- function(binning, data){
  lRandom <- lossCalc(data, binning=binning, type="random", newData = TRUE)
  lMean <- lossCalc(data, binning=binning, type="standard", newData=TRUE)
  datplot <- rbind(cbind(unique(lRandom$NewData), Algorithm="Random"), cbind(unique(lMean$NewData), Algorithm="Standard"))
  datplot$xbin <- round(datplot[,1]/binning[1],0)*binning[1]
  datplot$ybin <- round(datplot[,2]/binning[2],0)*binning[2]
  #datplot2 <- ddply(datplot, .(G, SO,Algorithm), summarize, count=length(G))
    qplot(xbin, ybin, geom="tile", data = datplot, fill = fsum , 
          ,main=paste("Baseball: Binning by (", binning[1], ", ", binning[2],")", sep="")) + facet_grid(.~Algorithm) +
        scale_fill_gradient(low="#56B1F7", high="#132B43", guide="legend", trans="log", breaks=c(1, 5, 50, 500, 5000)) + theme_bw() + theme(legend.position="bottom")
    ggsave(paste("./RandomStd", binning[1], binning[2], ".pdf", sep=""), width=8, height=5, units="in", dpi=1200)
  return(rbind(
    cbind(binX = binning[1], binY = binning[2], Function = "Random", lRandom$Loss), 
    cbind(binX = binning[1], binY = binning[2], Function = "Standard", lMean$Loss)))
}

@@

<<tablelossexample, echo=FALSE, eval=FALSE, results='asis',cache=FALSE>>=
loss0.1 <- lossCalc(data=X[,c(1:2)], binning=c(.1, .1), type="standard", newData=TRUE)
loss0.25 <- lossCalc(data=X[,c(1:2)], binning=c(.25, .25), type="standard", newData=TRUE)
loss0.5 <- lossCalc(data=X[,c(1:2)], binning=c(.5, .5), type="standard", newData=TRUE)
loss1 <- lossCalc(data=X[,c(1:2)], binning=c(1, 1), type="standard",newData=TRUE)

loss0.1 <- data.frame(with(loss0.1, c(Loss, rows=nrow(ddply(NewData, .(x,y), summarise, Freq=sum(fsum))), cols=ncol(ddply(NewData, .(x,y), summarise, Freq=sum(fsum))))))
loss0.25 <- data.frame(with(loss0.25, c(Loss, rows=nrow(ddply(NewData, .(x,y), summarise, Freq=sum(fsum))), cols=ncol(ddply(NewData, .(x,y), summarise, Freq=sum(fsum))))))
loss0.5 <- data.frame(with(loss0.5, c(Loss, rows=nrow(ddply(NewData, .(x,y), summarise, Freq=sum(fsum))), cols=ncol(ddply(NewData, .(x,y), summarise, Freq=sum(fsum))))))
loss1 <- data.frame(with(loss1, c(Loss, rows=nrow(ddply(NewData, .(x,y), summarise, Freq=sum(fsum))), cols=ncol(ddply(NewData, .(x,y), summarise, Freq=sum(fsum))))))

table.loss <- data.frame(rbind(c(binwidth=0, TotalLoss.x=0, TotalLoss.y=0, TotalLoss.LogFreq=0, rows=200), 
                               c(binwidth=.1, loss0.1[5:7]*100, loss0.1[8]), 
                               c(binwidth=.25, loss0.25[5:7]*100, loss0.25[8]), 
                               c(binwidth=.5, loss0.5[5:7]*100, loss0.5[8]), 
                               c(binwidth=1, loss1[5:7]*100, loss1[8])))
names(table.loss) <- c("Bin Width", "% Loss in X", "% Loss in Y", "% Loss in log(Freq)", "Size of Data")
print(xtable(table.loss, digits=c(0, 2, 4, 4, 4, 0), caption="Loss due to binning for all examples in figure 4.", label="tablelossexample"), include.rownames=FALSE, table.placement="hbtp")
@

@@

@@

\begin{figure}[hbtp]\centering
%\subfloat[Million-point sample of Airline Data]{\includegraphics[width=.44\linewidth, keepaspectratio=TRUE]{images/MillionSample.pdf}}\hfil
%   Swap Back at later date (shrinks size of pdf to managable)
\subfloat[Million-point sample of Airline Data. \ktm{This plot is not compiling and needs to be reconstructed (my apologies)}]{\includegraphics[width=.44\linewidth, keepaspectratio=TRUE]{images/RplotFiller.pdf}}\hfil
\subfloat[Minimal binning of Airline Data]{\includegraphics[width=.56\linewidth, keepaspectratio=TRUE]{images/AirlineStdBinning1min.pdf}}
\caption{\label{airline-scatter}Scheduled and actual departure times of  flights across the United States in 2011. The plot on the left is based on a sample of the data, the plot on the right shows all flights. The big patterns are visible in both plots, but the plot on the left misses some of the finer level details in scheduling that is visible in the plot on the right.}
\end{figure}

@@

<<airlinesetup,echo=FALSE, eval=FALSE>>=
setwd("./data/")
# airline <- do.call("rbind", lapply(list.files(), function(i) read.csv(i)))
# write.csv(airline, "airline.csv", row.names=FALSE)
airline <- read.csv("airline.csv")
na.departure <- (is.na(airline[,8]) + is.na(airline[,9]))> 0
airline <- airline[!na.departure,8:9]

@@

airline.freq <- ddply(airline, .(CRS_DEP_TIME, DEP_TIME), function(i) cbind(unique(i), Freq=nrow(i)))
write.csv(airline.freq, "ReducedAirlineData.csv", row.names=FALSE)
airline.freq <- read.csv("ReducedAirlineData.csv")
airline.freq.plot <- airline.freq
airline.freq.plot[,1] <- convert.time(airline.freq.plot[,1])
airline.freq.plot[,2] <- convert.time(airline.freq.plot[,2])
airline.freq.plot <- airline.freq.plot[order(airline.freq.plot$Freq),]
airline.dummy <- airline.freq.plot[1,]
airline.dummy[1,1] <- NA
qplot(data=airline.freq.plot, x=CRS_DEP_TIME, y=DEP_TIME, fill=Freq, asp=1, geom="blank", xlab="Scheduled Departure Time", ylab="Actual Departure Time", main="Airline Departure Times (1 minute bins)") + geom_rect(aes(xmin=CRS_DEP_TIME-3/60, xmax=CRS_DEP_TIME+3/60, ymin=DEP_TIME-3/60, ymax=DEP_TIME+3/60, fill=Freq), alpha=I(.5)) + geom_rect(data=airline.dummy, aes(xmin=CRS_DEP_TIME-3/60, xmax=CRS_DEP_TIME+3/60, ymin=DEP_TIME-3/60, ymax=DEP_TIME+3/60, fill=Freq), aes.inherit=FALSE) + theme_bw()+ theme(legend.position="right") +scale_fill_gradientn(colours=c("#56B1F7", "#132B43"), guide="legend", trans="log", breaks=c(1, 10, 100, 1000, 10000))+ scale_x_continuous(breaks=c(0, 6, 12, 18, 24)) + scale_y_continuous(breaks=c(0, 6, 12, 18, 24))
ggsave("AirlineStdBinning1min.pdf", width=7.5, height=6, dpi=2304)

@@

air.binned5 <- binStd(airline.freq, c(5, 5))
air.binned5 <- ddply(air.binned5, .(CRS_DEP_TIME, DEP_TIME), function(i) cbind(convert.time(unique(i[,1:2])), Freq=sum(i$Freq)))
write.csv(air.binned5, "airlineBin5Std.csv", row.names=FALSE)

# air.binned5.rdm <- binRdm(cbind(airline, Freq=1), c(5, 5))
# air.binned5.rdm2 <- ddply(air.binned5.rdm, .(CRS_DEP_TIME, DEP_TIME), function(i) cbind(convert.time(unique(i[,1:2]), Freq=sum(i$Freq)))
# write.csv(air.binned5.rdm2, "airlineBin5Rdm.csv", row.names=FALSE)

>>>>

@@

      temp <- data.frame(   
                x=c(-20, -10,   0,  10,  20,  30,  40, -20, -20, -20, -20, -20, -20), 
             xend=c(-20, -10,   0,  10,  20,  30,  40,  40,  40,  40,  40,  40,  40), 
                y=c(  0,   0,   0,   0,   0,   0,   0,   0,  10,  20,  30,  40,  50), 
             yend=c( 50,  50,  50,  50,  50,  50,  50,   0,  10,  20,  30,  40,  50), 
            group=c(  1,   2,   3,   4,   5,   6,   7,   8,   9,  10,  11,  12,  13))
      qplot(data=d[,], x=x, y=y, alpha=a,  geom="point", 
            xlim=c(-20, 40), ylim=c(0, 50)) +
        annotate("polygon", x=c(10, 10, 20, 20), y=c(20, 30, 30, 20), 
                 fill="blue", alpha=.25) +
        annotate("polygon", x=c(0, 0, 10, 10), y=c(20, 30, 30, 20), 
                 fill="green", alpha=.25) +
        scale_alpha(range=c(.5, 1), guide=FALSE) +
      #  scale_colour_identity(guide=FALSE) + 
      #  scale_shape(solid=T,guide=FALSE) +
        theme_bw() +
        annotate("point", x=rep(seq(-15, 35, 10),each=7), 
               y=rep(seq(-5, 55, 10), times=6), 
               shape='+', size=8, colour="gray90") + 
        geom_point(data=d[which(!is.na(d$xbin)),], 
                   aes(x=x, y=y, shape=factor(xbin), colour=factor(xbin)), 
                   size=I(3), aes.inherit=FALSE) + 
        scale_shape_manual(name="Bin", values=c("5"=16, "15"=15), 
                           labels=c("(5,25)", "(15, 25)"),guide=FALSE) +
        scale_colour_manual(values=c("5"="green", "15"="blue"), name="Bin", 
                            labels=c("(5,25)", "(15, 25)"),guide=FALSE)+
        annotate("segment", x=temp$x, y=temp$y, 
                 xend=temp$xend, yend=temp$yend, group=temp$group) # + ylim(c(0, 50))
      ggsave("./figure/Binning-DataAndBin.pdf", width=5, height=5, units="in")

@@

\begin{figure}[hbtp]
  \centering
  \subfloat[Standard Algorithm]{\includegraphics[keepaspectratio=true, width=3in]{./figure/Binning-DataAndBin.pdf}}
%   \subfloat{\includegraphics[keepaspectratio=TRUE, width=2in]
%   {./figure/RandomBinningDataPoints1}}
  \subfloat[Random Algorithm]{\includegraphics[keepaspectratio=TRUE, width=3in]
  {./figure/RandomBinningDataPoints2.pdf}}
  \caption{Points are assigned to bins centered at (5, 25) an (15,25). Points in the standard algorithm are contained within the defined bin region, where points from the random algorith may be assigned to bins from outside the defined bin regions.}
  \label{fig:BinningTutorial}
\end{figure}

@@

\section{Discussion and Examples}
\subsection{Comparison of Loss in Binning Algorithms}
<<LossTableGraphs,echo=F,results="hide",cache=TRUE, eval=TRUE>>=
# bins <- rbind(c(1,1), c(2,2), c(2,5), c(5, 2), c(4,4), c(3, 5), c(5, 3), c(5, 5), c(5, 25), c(25, 5), c(2, 50), c(50, 2), c(5, 50))
computeLoss <- function(binning, data){
  lRandom <- lossCalc(data, binning=binning, type="random", newData = TRUE)
  lMean <- lossCalc(data, binning=binning, type="standard", newData=TRUE)
  datplot <- rbind(cbind(unique(lRandom$NewData), Algorithm="Random"), cbind(unique(lMean$NewData), Algorithm="Standard"))
  datplot$xbin <- round(datplot[,1]/binning[1],0)*binning[1]
  datplot$ybin <- round(datplot[,2]/binning[2],0)*binning[2]
  #datplot2 <- ddply(datplot, .(G, SO,Algorithm), summarize, count=length(G))
    qplot(xbin, ybin, geom="tile", data = datplot, fill = Freq , 
          ,main=paste("Baseball: Binning by (", binning[1], ", ", binning[2],")", sep="")) + facet_grid(.~Algorithm) +
        scale_fill_gradient(low="#56B1F7", high="#132B43", guide="legend", trans="log", breaks=c(1, 5, 50, 500, 5000)) + theme_bw() + theme(legend.position="bottom")
    ggsave(paste("./RandomStd", binning[1], binning[2], ".pdf", sep=""), width=8, height=5, units="in", dpi=1200)
  return(rbind(
    cbind(binX = binning[1], binY = binning[2], Function = "Random", lRandom$Loss), 
    cbind(binX = binning[1], binY = binning[2], Function = "Standard", lMean$Loss)))
}

@@

\hh{the  code following this comment in the Rnw file is, I think, broken beyond repair - but we should include a table like the one it produced in the paper.  - we will have different losses to report, though}

% 
% <<tablelossexample, echo=FALSE, eval=FALSE, results='asis',cache=FALSE>>=
% loss0.1 <- lossCalc(data=X[,c(1:2)], binning=c(.1, .1), type="standard", newData=TRUE)
% loss0.25 <- lossCalc(data=X[,c(1:2)], binning=c(.25, .25), type="standard", newData=TRUE)
% loss0.5 <- lossCalc(data=X[,c(1:2)], binning=c(.5, .5), type="standard", newData=TRUE)
% loss1 <- lossCalc(data=X[,c(1:2)], binning=c(1, 1), type="standard",newData=TRUE)
% 
% loss0.1 <- data.frame(with(loss0.1, c(Loss, rows=nrow(ddply(NewData, .(x,y), summarise, Freq=sum(Freq))), cols=ncol(ddply(NewData, .(x,y), summarise, Freq=sum(Freq))))))
% loss0.25 <- data.frame(with(loss0.25, c(Loss, rows=nrow(ddply(NewData, .(x,y), summarise, Freq=sum(Freq))), cols=ncol(ddply(NewData, .(x,y), summarise, Freq=sum(Freq))))))
% loss0.5 <- data.frame(with(loss0.5, c(Loss, rows=nrow(ddply(NewData, .(x,y), summarise, Freq=sum(Freq))), cols=ncol(ddply(NewData, .(x,y), summarise, Freq=sum(Freq))))))
% loss1 <- data.frame(with(loss1, c(Loss, rows=nrow(ddply(NewData, .(x,y), summarise, Freq=sum(Freq))), cols=ncol(ddply(NewData, .(x,y), summarise, Freq=sum(Freq))))))
% 
% table.loss <- data.frame(rbind(c(binwidth=0, TotalLoss.x=0, TotalLoss.y=0, TotalLoss.LogFreq=0, rows=200), 
%                                c(binwidth=.1, loss0.1[5:7]*100, loss0.1[8]), 
%                                c(binwidth=.25, loss0.25[5:7]*100, loss0.25[8]), 
%                                c(binwidth=.5, loss0.5[5:7]*100, loss0.5[8]), 
%                                c(binwidth=1, loss1[5:7]*100, loss1[8])))
% names(table.loss) <- c("Bin Width", "% Loss in X", "% Loss in Y", "% Loss in log(Freq)", "Size of Data")
% print(xtable(table.loss, digits=c(0, 2, 4, 4, 4, 0), caption="Loss due to binning for all examples in figure 4.", label="tablelossexample"), include.rownames=FALSE, table.placement="hbtp")
% @

@@

\hh{This is a valid concern, but we have to fix the notation at the beginning first, before we can address it.}

@@

\begin{figure}[hbtp]\centering
\subfloat[Million-point sample of Airline Data]{\includegraphics[width=.44\linewidth, keepaspectratio=TRUE]{images/MillionSample.pdf}}\hfill
%   Swap Back at later date (shrinks size of pdf to managable)
%\subfloat[Million-point sample of Airline Data. \ktm{This plot is not compiling and needs to be reconstructed (my apologies)}]{\includegraphics[width=.44\linewidth, keepaspectratio=TRUE]{images/RplotFiller.pdf}}\hfil
\subfloat[Minimal binning of Airline Data]{\includegraphics[width=.56\linewidth, keepaspectratio=TRUE]{images/AirlineStdBinning1min.pdf}}
\caption{\label{airline-scatter}Scheduled and actual departure times of  flights across the United States in 2011. The plot on the left is based on a sample of the data, the plot on the right shows all flights. The big patterns are visible in both plots, but the plot on the left misses some of the finer level details in scheduling that is visible in the plot on the right.}
\end{figure}

@@

<<airlinesetup,echo=FALSE, eval=FALSE>>=
setwd("./data/")
if (!file.exists("airline.csv.gz")) {
  # HH: I'm not sure, which files these should be ... 
  # airline <- do.call("rbind", lapply(list.files(), function(i) read.csv(i)))

  # write.csv(airline, "airline.csv", row.names=FALSE)
}
airline <- read.csv("airline.csv.gz")
na.departure <- (is.na(airline[,8]) + is.na(airline[,9]))> 0
airline <- airline[!na.departure,8:9]

@@

if (!file.exists("ReducedAirlineData.csv")) {
  airline.freq <- ddply(airline, .(CRS_DEP_TIME, DEP_TIME), function(i) cbind(unique(i), Freq=nrow(i)))
  write.csv(airline.freq, "ReducedAirlineData.csv", row.names=FALSE)
}
airline.freq <- read.csv("ReducedAirlineData.csv")
airline.freq.plot <- airline.freq
airline.freq.plot[,1] <- convert.time(airline.freq.plot[,1])
airline.freq.plot[,2] <- convert.time(airline.freq.plot[,2])
airline.freq.plot <- airline.freq.plot[order(airline.freq.plot$Freq),]
airline.dummy <- airline.freq.plot[1,]
airline.dummy[1,1] <- NA
qplot(data=airline.freq.plot, x=CRS_DEP_TIME, y=DEP_TIME, fill=Freq, asp=1, geom="blank", xlab="Scheduled Departure Time", ylab="Actual Departure Time", main="Airline Departure Times (1 minute bins)") + geom_rect(aes(xmin=CRS_DEP_TIME-3/60, xmax=CRS_DEP_TIME+3/60, ymin=DEP_TIME-3/60, ymax=DEP_TIME+3/60, fill=Freq), alpha=I(.5)) + geom_rect(data=airline.dummy, aes(xmin=CRS_DEP_TIME-3/60, xmax=CRS_DEP_TIME+3/60, ymin=DEP_TIME-3/60, ymax=DEP_TIME+3/60, fill=Freq), aes.inherit=FALSE) + theme_bw()+ theme(legend.position="right") +scale_fill_gradientn(colours=c("#56B1F7", "#132B43"), guide="legend", trans="log", breaks=c(1, 10, 100, 1000, 10000))+ scale_x_continuous(breaks=c(0, 6, 12, 18, 24)) + scale_y_continuous(breaks=c(0, 6, 12, 18, 24))
ggsave("AirlineStdBinning1min.pdf", width=7.5, height=6, dpi=2304)

@@

if (!file.exists("airlineBin5Std.csv")) {
  air.binned5 <- binStd(airline.freq, c(5, 5))
  air.binned5 <- ddply(air.binned5, .(CRS_DEP_TIME, DEP_TIME), function(i) cbind(convert.time(unique(i[,1:2])), Freq=sum(i$Freq)))
  write.csv(air.binned5, "airlineBin5Std.csv", row.names=FALSE)
}
# air.binned5.rdm <- binRdm(cbind(airline, Freq=1), c(5, 5))
# air.binned5.rdm2 <- ddply(air.binned5.rdm, .(CRS_DEP_TIME, DEP_TIME), function(i) cbind(convert.time(unique(i[,1:2]), Freq=sum(i$Freq)))
# write.csv(air.binned5.rdm2, "airlineBin5Rdm.csv", row.names=FALSE)
